Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective

### 2025-06-03 MS2_Krill.Rmd
adding scenarios

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@REF]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@REF], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@REF]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

@@

## Krill Length-to-Age Model Description

@@

ft_1 <- autofit(ft_env)
ft_2 <- set_table_properties(ft_1, width = .5, layout = "autofit")
ft_2

@@

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [REF] and SS3 addresses these approaches. According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment. This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;

@@

The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population (Figure \@ref(fig:pearson)). In scenario s1.1, substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s. Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects. s1.2  shows some improvement in model fit, with slightly more balanced residuals. However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition. In s1.3, there is a reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions. This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins. Finally, s1.4 shows the most consistent reduction in Pearson residuals across time and size structure. The residual patterns also indicate potentially concerning model misfits, particularly among adult individuals in several survey fleets operating in the Bransfield Strait, as well as among the largest size classes in the predator fleet, which likely represents penguin consumption data. The residuals exhibit consistent over- or under-estimation in specific length bins over multiple years, suggesting that the current model structure may not fully capture key processes affecting the abundance or selectivity of adult individuals. In the case of surveys in Bransfield, this could reflect spatial mismatches between predicted and observed size distributions or biases in survey selectivity for larger size classes. This spatial heterogeneity aligns with previous findings identifying this area as a core juvenile krill habitat [@Mardones2025; @Veytia2021, Vetia et al., 2019, REF, REF]. Similarly, the residuals associated with the predator fleet suggest that the model may inadequately represent size-selective predation or the temporal dynamics of predator consumption. These patterns warrant further investigation into the assumptions underlying length selectivity and mortality, particularly for the adult component of the population, as well as potential ecosystem-driven variability not currently integrated in the model.

@@

(FINAL TABLE Suggested by MJ)

@@

@@

# Data Availability

Data will be made available on request.

>>>>

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

@@

## Length-to-Age Model Description

@@

ft_1 <- autofit(ft_env)
ft_2 <- set_table_properties(ft_1, width = .5, layout = "autofit")
ft_2
```

@@

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [@Marshall2019] and SS3 addresses these approaches. According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment. This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;

@@

The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population (Figure \@ref(fig:pearson)). In scenario s1.1, substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s. Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects. s1.2  shows some improvement in model fit, with slightly more balanced residuals. However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition. In s1.3, there is a reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions. This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins. Finally, s1.4 shows the most consistent reduction in Pearson residuals across time and size structure. The residual patterns also indicate potentially concerning model misfits, particularly among adult individuals in several survey fleets operating in the Bransfield Strait, as well as among the largest size classes in the predator fleet, which likely represents penguin consumption data. The residuals exhibit consistent over- or under-estimation in specific length bins over multiple years, suggesting that the current model structure may not fully capture key processes affecting the abundance or selectivity of adult individuals. In the case of surveys in Bransfield, this could reflect spatial mismatches between predicted and observed size distributions or biases in survey selectivity for larger size classes. This spatial heterogeneity aligns with previous findings identifying this area as a core juvenile krill habitat [@Mardones2025; @Veytia2021, @Perry2020: @Perry2019]. Similarly, the residuals associated with the predator fleet suggest that the model may inadequately represent size-selective predation or the temporal dynamics of predator consumption. These patterns warrant further investigation into the assumptions underlying length selectivity and mortality, particularly for the adult component of the population, as well as potential ecosystem-driven variability not currently integrated in the model.

@@

<!-- (FINAL TABLE Suggested by MJ) -->

@@

<!-- Because M2 is a MGparm, it can be time-varying like any other MGparm. This is -->
<!-- important because M2, as a component added to base M, will probably always need to be -->
<!-- time-varying by blocks, random walk or linkage to external driver. (SS3 manual) -->

@@

<!-- # Data Availability -->

<!-- Data will be made available on request. -->
